msc dissertation

### 2021-08-13 msc dissertation code.Rmd
latest

---
title: "msc dissertation"
author: "Halleyson Li"
date: "'r Sys.Date()'"
output: html_document
line-citations: apalike
linkcolor: blue
frontsize: 12pt
---

@@

@@

pennycook2 <- read_csv("Pennycook et al._Study 2.csv")
# Quick check
head(pennycook2)

@@

@@

#Choose with the column position 
library("dplyr")
pennystudy2r1 <- pennycook2 %>% 
  select (c(1, 210, 211, 212, 213, 214, 215, 216, 217, 218, 219, 220, 221, 222, 223, 224, 225, 226, 227, 228, 229, 230, 231, 232, 233, 234, 235, 236, 237, 238, 239, 2, 3, 4, 57, 60, 250, 252, 284,))

#re-code sharing intentions (fake/real) using the scale midpoint (1-3 = 0, 4-6 = 1)

@@

###find out the number of participants (obs) for each attention check level
attention1 <- subset(pennys2r1long_nona, attentioncheck %in% c("1","2","3"))
attention2 <- subset(pennys2r1long_nona, attentioncheck %in% c("2","3"))
attention3 <- subset(pennys2r1long_nona, attentioncheck %in% c("3"))

@@

## relase I should not dichotomise the scale as it loses info and reduces the power.
```{r, message=FALSE}
#recreate a data frame that does not rescale the data to either 0,1 but from 0 to 1,so divided each value by rating = (rating - 1)/5

pennystudy2r1nodi <- pennycook2 %>% 
  select (c(1, 210, 211, 212, 213, 214, 215, 216, 217, 218, 219, 220, 221, 222, 223, 224, 225, 226, 227, 228, 229, 230, 231, 232, 233, 234, 235, 236, 237, 238, 239, 2, 3, 4, 57, 60, 250, 252, 284,))

@@

##no idea why it doesn't show conditions (control vs treatment) so I merge original condition to the new one

@@

##create a new column that recode item number from fake and real to 1-30, with 1-15 indicate each fake items, 16-30 indicate each real items

pennys2r1longnodi_nona <- within(pennys2r1longnodi_nona,{
  item_id <- NA
  item_id[newstype == "Fake1_1"] <- 1
  item_id[newstype == "Fake1_2"] <- 2
  item_id[newstype == "Fake1_3"] <- 3
  item_id[newstype == "Fake1_4"] <- 4
  item_id[newstype == "Fake1_5"] <- 5
  item_id[newstype == "Fake1_6"] <- 6
  item_id[newstype == "Fake1_7"] <- 7
  item_id[newstype == "Fake1_8"] <- 8
  item_id[newstype == "Fake1_9"] <- 9
  item_id[newstype == "Fake1_10"] <- 10
  item_id[newstype == "Fake1_11"] <- 11
  item_id[newstype == "Fake1_12"] <- 12
  item_id[newstype == "Fake1_13"] <- 13
  item_id[newstype == "Fake1_14"] <- 14
  item_id[newstype == "Fake1_15"] <- 15
  item_id[newstype == "Real1_1"] <- 16
  item_id[newstype == "Real1_2"] <- 17
  item_id[newstype == "Real1_3"] <- 18
  item_id[newstype == "Real1_4"] <- 19
  item_id[newstype == "Real1_5"] <- 20
  item_id[newstype == "Real1_6"] <- 21
  item_id[newstype == "Real1_7"] <- 22
  item_id[newstype == "Real1_8"] <- 23
  item_id[newstype == "Real1_9"] <- 24
  item_id[newstype == "Real1_10"] <- 25
  item_id[newstype == "Real1_11"] <- 26
  item_id[newstype == "Real1_12"] <- 27
  item_id[newstype == "Real1_13"] <- 28
  item_id[newstype == "Real1_14"] <- 29
  item_id[newstype == "Real1_15"] <- 30
})

@@

attention1nodi <- subset(pennys2r1longnodi_nona, attentioncheck %in% c("1","2","3"))
attention2nodi <- subset(pennys2r1longnodi_nona, attentioncheck %in% c("2","3"))
attention3nodi <- subset(pennys2r1longnodi_nona, attentioncheck %in% c("3"))

@@

install.packages("effsize")
library(effsize)

@@

sapply(pennys2r1longnodi_nona,mode)

pennys2nodicohenc <- subset(pennys2r1longnodi_nona, condition %in% c("0"))
pennys2nodicohent <- subset(pennys2r1longnodi_nona, condition %in% c("1"))

#condition vs measurement
cohen.d(pennys2nodicohent$Real,pennys2nodicohent$Fake, na.rm = TRUE)
cohen.d(pennys2nodicohenc$Real,pennys2nodicohenc$Fake, na.rm = TRUE)

@@

## reproduction of figure 2 without considering attention check

@@

library(ggplot2)

ggsave("All Subjects bar chart.png")
```

##Try something new, see if it makes the results better
###GLM
```{r, message=FALSE}
## use glm, stepwise approach to see how each factor explaining the sharing intention. The family for the model is binomial since the dependent variable is dichotomous.
glm0 <- glm(measurement ~ 1, data = pennys2r1long_nona, family = binomial(link = "logit"))

###Ordinal logistic regression
```{r, message=FALSE}
#Ordinal logistic regression

install.packages("ordinal")
install.packages("ggeffects")
install.packages("ggrepel")

@@

ggplot(ggpredict_ordinaldata3, aes(x = Rating, y = predicted)) + 
  geom_point(aes(color = Headlineveracity), position =position_dodge(width = 0.5)) + geom_errorbar(aes(ymin = conf.low, ymax = conf.high, color = Headlineveracity), position = position_dodge(width = 0.5), width = 0.3) + 
  theme_minimal() + facet_wrap(~group) + xlab("Likelihood of Sharing") + ylab("Predicted Probability of Sharing") + geom_label_repel(aes(label=sprintf("%.3f", predicted),fill = Headlineveracity), box.padding = 0.5, min.segment.length = unit(0, 'lines'), size = 3) + labs(title = "All 3 Screener Correct")

@@

### Tests (irrelevant)

#z scored
scale(pennystudy2r1$D)

@@

### graph 3 attention check above and equal to 2
```{r, message=FALSE}
#condition 3: above and equal to 2
pennystudy2plot3 <- pennys2r1longa2 %>%
  select (c(1, 2, 3, 4))

@@

#plot the graph under the proportion of overall observation (856)
ggplot(p3, aes(x = condition, y = as.numeric(percent), fill = newstype)) + 
  geom_bar(stat = 'identity',position = 'dodge') + 
  geom_text(
    aes(label = as.numeric (percent)), vjust = 1.5, position = position_dodge(0.9))
```
### graph 3 attention check above and equal to 2: mean
```{r, message=FALSE}
#condition 3: above and equal to 2
pennystudy2plot3 <- pennys2r1longa2 %>%
  select (c(1, 2, 3, 4))

@@

### graph 4 attention check equal to 3
```{r, message=FALSE}
#condition 3: equal to 3
pennystudy2plot4 <- pennys2r1longa3 %>%
  select (c(1, 2, 3, 4))

>>>>

---
title: "msc dissertation"
author: "Halleyson Li 200256823"
date: '''r Sys.Date()'''
output:
  html_document: default
  pdf_document: default
line-citations: apalike
linkcolor: blue
frontsize: 12pt
---

@@

Because the original data was not encoded in UFT-8, so after download the original data, has to open it with excel and re-save it with UFT-8 encoded .csv file. But if knit/publish is not required, then this process is not required, but for loading the data (using readr), it has to be read_csv but not read.csv.

@@

pennycook2 <- read_csv("Pennycooketal Study 2.csv")

# Quick check
head(pennycook2)
```

@@

```{r, message=FALSE, results='hide'}

@@

#Choose with the column position 
library("dplyr")
pennystudy2r1 <- pennycook2 %>% 
  dplyr::select(c(1, 210, 211, 212, 213, 214, 215, 216, 217, 218, 219, 220, 221, 222, 223, 224, 225, 226, 227, 228, 229, 230, 231, 232, 233, 234, 235, 236, 237, 238, 239, 2, 3, 4, 57, 60, 250, 252, 284,))

#re-code sharing intentions (fake/real) using the scale midpoint (1-3 = 0, 4-6 = 1)
###(06/08/21 update: realize I could do it after change it from wide to long, then only the column: "measurement" needs to perform the change.)

@@

###find out the number of participants (obs) for each attention check level (use it to compare with No of participants in Table S5).

No.participantsa1 <- subset(pennystudy2r1, dummycheck %in% c("1","2","3"))
No.participantsa2 <- subset(pennystudy2r1, dummycheck %in% c("2","3"))
No.participantsa3 <- subset(pennystudy2r1, dummycheck %in% c("3"))

@@

## relase I should not dichotomise the scale as it loses info and reduces the power.
```{r, message=FALSE, results='hide'}
#recreate a data frame that does not rescale the data to either 0,1 but from 0 to 1,so divided each value by rating = (rating - 1)/5

pennystudy2r1nodi <- pennycook2 %>% 
 dplyr::select (c(1, 210, 211, 212, 213, 214, 215, 216, 217, 218, 219, 220, 221, 222, 223, 224, 225, 226, 227, 228, 229, 230, 231, 232, 233, 234, 235, 236, 237, 238, 239, 2, 3, 4, 57, 60, 250, 252, 284,))

@@

## I merge original condition to the new one

@@

@@

##create a new column that recode item number from fake and real to 1-30, with 1-15 indicate each fake items, 16-30 indicate each real items, need it for clustering robust standard errors.

pennys2r1longnodi_nona <- within(pennys2r1longnodi_nona,{
  item_id <- NA
  item_id[newstype == "Fake1_1"] <- 1
  item_id[newstype == "Fake1_2"] <- 2
  item_id[newstype == "Fake1_3"] <- 3
  item_id[newstype == "Fake1_4"] <- 4
  item_id[newstype == "Fake1_5"] <- 5
  item_id[newstype == "Fake1_6"] <- 6
  item_id[newstype == "Fake1_7"] <- 7
  item_id[newstype == "Fake1_8"] <- 8
  item_id[newstype == "Fake1_9"] <- 9
  item_id[newstype == "Fake1_10"] <- 10
  item_id[newstype == "Fake1_11"] <- 11
  item_id[newstype == "Fake1_12"] <- 12
  item_id[newstype == "Fake1_13"] <- 13
  item_id[newstype == "Fake1_14"] <- 14
  item_id[newstype == "Fake1_15"] <- 15
  item_id[newstype == "Real1_1"] <- 16
  item_id[newstype == "Real1_2"] <- 17
  item_id[newstype == "Real1_3"] <- 18
  item_id[newstype == "Real1_4"] <- 19
  item_id[newstype == "Real1_5"] <- 20
  item_id[newstype == "Real1_6"] <- 21
  item_id[newstype == "Real1_7"] <- 22
  item_id[newstype == "Real1_8"] <- 23
  item_id[newstype == "Real1_9"] <- 24
  item_id[newstype == "Real1_10"] <- 25
  item_id[newstype == "Real1_11"] <- 26
  item_id[newstype == "Real1_12"] <- 27
  item_id[newstype == "Real1_13"] <- 28
  item_id[newstype == "Real1_14"] <- 29
  item_id[newstype == "Real1_15"] <- 30})

##subset based on attetnion check
attention1nodi <- subset(pennys2r1longnodi_nona, attentioncheck %in% c("1","2","3"))
attention2nodi <- subset(pennys2r1longnodi_nona, attentioncheck %in% c("2","3"))
attention3nodi <- subset(pennys2r1longnodi_nona, attentioncheck %in% c("3"))

@@

library(effsize)

@@

#check the class
sapply(pennys2r1longnodi_nona,mode)

#subset the table based on condition
pennys2nodicohenc <- subset(pennys2r1longnodi_nona, condition %in% c("0"))
pennys2nodicohent <- subset(pennys2r1longnodi_nona, condition %in% c("1"))

#condition vs measurement effect size, using participants' average measurement.
cohen.d(pennys2nodicohent$Real,pennys2nodicohent$Fake, na.rm = TRUE)
cohen.d(pennys2nodicohenc$Real,pennys2nodicohenc$Fake, na.rm = TRUE)

@@

## reproduction of figure 2 (without considering attention check)

@@

library(ggplot2)
library(DescTools)

#create a table contain mean and CI for each headline veracity in each condition
z <- with(pennys2r1long_nona, 
          aggregate(measurement, list(condition, headlineveracity), MeanCI))
z <- xtabs(x ~ Group.1 + Group.2, z)

z = data.frame(z)

#check the results with equation I know for sure
dummy <- aggregate(measurement~condition+headlineveracity, data = pennys2r1long_nona, FUN = mean)

#reshape the graph
z1 <- subset(z, Var3 %in% c("mean"))
z2 <- subset(z, Var3 %in% c("lwr.ci"))
z3 <- subset(z, Var3 %in% c("upr.ci"))

zsummary <- cbind(z1,z2$Freq,z3$Freq)

zsummary <- zsummary %>% 
  dplyr::select (c(1, 2, 4, 5, 6))

names(zsummary)[1] <- "condition"
names(zsummary)[2] <- "headlineveracity"
names(zsummary)[3] <- "percent"
names(zsummary)[4] <- "lowci"
names(zsummary)[5] <- "upci"

#change the value to %
zsummary$percent <- zsummary$percent*100
zsummary$lowci <- zsummary$lowci*100
zsummary$upci <- zsummary$upci*100
str(zsummary)

#plot the graph with error bar indicate CI

ggplot(zsummary, aes(x = condition, y = percent, fill = forcats::fct_rev(headlineveracity))) + 
  geom_bar(stat = 'identity',position = 'dodge',width = 0.5) + 
  geom_text(
    aes(label = sprintf("%.2f",percent)), vjust = 2.5, position = position_dodge(0.5)) +
  xlab("Condition") + 
  ylab("Likelihood of Sharing (%)") +
  scale_x_discrete(labels= c("0"="Control","1"="Treatment"))+
  scale_fill_manual(values = c("orange2","forestgreen"), name = "Headline Veracity", labels = c("True", "False")) + labs(title = "All Subjects") + 
  geom_errorbar(aes(ymin = lowci, ymax = upci,width = 0.2),position = position_dodge(0.5), size = 0.6)

##tips: forcats::fct_rev() is to reverse the order, use scale_fill_manual(values = to change colour

ggsave("All Subjects bar chart.png")

##Try something new, see if it makes the results better

###Ordinal logistic regression (best to apply when using likert scale)
```{r, message=FALSE}
#Ordinal logistic regression

@@

ggplot(ggpredict_ordinaldata3, aes(x = Rating, y = predicted)) + 
  geom_point(aes(color = Headlineveracity), position =position_dodge(width = 0.5)) + geom_errorbar(aes(ymin = conf.low, ymax = conf.high, color = Headlineveracity), position = position_dodge(width = 0.5), width = 0.3) + 
  theme_minimal() + facet_wrap(~group) + xlab("Likelihood of Sharing") + ylab("Predicted Probability of Sharing") + geom_label_repel(aes(label=sprintf("%.3f", predicted),fill = Headlineveracity), box.padding = 0.55, min.segment.length = unit(0, 'lines'), size = 2.5) + labs(title = "All 3 Screener Correct")

@@

### code testing (irrelevant)

###GLM (bad idea, because it dicotomize the results leads to loss information and verlidity)

## use glm, stepwise approach to see how each factor explaining the sharing intention. The family for the model is binomial since the dependent variable is dichotomous.
glm0 <- glm(measurement ~ 1, data = pennys2r1long_nona, family = binomial(link = "logit"))

##subset the attention based on attention check
attention1 <- subset(pennys2r1long_nona, attentioncheck %in% c("1","2","3"))
attention2 <- subset(pennys2r1long_nona, attentioncheck %in% c("2","3"))
attention3 <- subset(pennys2r1long_nona, attentioncheck %in% c("3"))

#old graph

```
#z scored
scale(pennystudy2r1$D)

@@

### graph 3 attention check above and equal to 2

#condition 3: above and equal to 2
pennystudy2plot3 <- pennys2r1longa2 %>%
  select (c(1, 2, 3, 4))

@@

#plot the graph under the proportion of overall observation (856)
ggplot(p3, aes(x = condition, y = as.numeric(percent), fill = newstype)) + 
  geom_bar(stat = 'identity',position = 'dodge') + 
  geom_text(
    aes(label = as.numeric (percent)), vjust = 1.5, position = position_dodge(0.9))
```
### graph 3 attention check above and equal to 2: mean

#condition 3: above and equal to 2
pennystudy2plot3 <- pennys2r1longa2 %>%
  select (c(1, 2, 3, 4))

@@

### graph 4 attention check equal to 3

#condition 3: equal to 3
pennystudy2plot4 <- pennys2r1longa3 %>%
  select (c(1, 2, 3, 4))
